Zero-shot task adaptation by meta-mapping from task to task

### 2019-09-04 writing/embedded_meta_learning_main.tex
Title + framing shift

\begin{document}
\title{Embedded Meta-Learning: Toward more flexible deep-learning models}
\author{%
Andrew K. Lampinen\\
Department of Psychology\\
Stanford University\\
\texttt{lampinen@stanford.edu}\\
\And
James L. McClelland\\
Department of Psychology\\
Stanford University\\
\texttt{mcclelland@stanford.edu}\\
}
\date{}
\maketitle

\begin{abstract}
How can deep learning systems flexibly reuse their knowledge? Toward this goal, we propose a new class of challenges, and a class of architectures that can solve them. The challenges are meta-mappings, which involve systematically transforming task behaviors to adapt to new tasks zero-shot. We suggest that the key to achieving these challenges is representing the task being performed along with the computations used to perform it. We therefore draw inspiration from meta-learning and functional programming to propose a class of Embedded Meta-Learning (EML) architectures that represent both data and tasks in a shared latent space. EML architectures are applicable to any type of machine learning task, including supervised learning and reinforcement learning. We demonstrate the flexibility of these architectures by showing that they can perform meta-mappings, i.e. that they can exhibit zero-shot remapping of behavior to adapt to new tasks, and explore several extensions. 
\end{abstract}

\vspace{-0.75em} % section
\section{Introduction}
\vspace{-0.5em} % section
Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep[e.g.][]{Lake2016, Marcus2018}. One fundamental reason for this is that humans are aware of what we are trying to compute and why. By contrast, there is a fundamental separation of knowledge within most deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between data and tasks. That is, a neural network's knowledge about what is being computed is only implicitly accessible to those computations. \par
\looseness=-1
There are a number of advantages to representing knowledge about data and tasks together. In particular, it can grant the ability to rapidly adapt behavior to a new task. The problem of rapid learning has been partially addressed by meta-learning systems \citep[e.g.][see also section \ref{sec_discussion}]{Santoro2016, Finn2017a, Finn2018, Stadie2018, Botvinick2019}. However, humans can use our knowledge of a task to flexibly alter our behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance without any retraining (i.e. zero-shot). Deep learning systems at present generally lack this representational flexibility. \par
In this paper, we propose a new class of tasks based on this idea: meta-mappings, i.e. mappings between tasks (see below). This type of transfer is easily accessible to humans \citep{Lake2016}, but is generally inaccessible to most deep-learning models. To address this challenge, we propose a new class of architectures which essentially take a functional perspective on meta-learning, and exploit the idea of homoiconicity. (A homoiconic programming language is one in which programs in the language can be manipulated by programs in the language, just as data can.) By treating both data and task behaviors as functions, we can conceptually think of both data \emph{and} learned task behaviors as transformable. This yields the ability to not only learn to solve new tasks, but to learn how to transform these solutions in response to changing task demands. We demonstrate that our architectures can flexibly remap their behavior to address the meta-mapping challenge. By allowing the network to recurse on its representations, our approach is able to achieve this flexibility parsimoniously, without adding additional parameters beyond what we use for basic meta-learning. Furthermore, the architecture may also offer other benefits, such as a useful framework for continual learning. We suggest that approaches like EML will be key to building more intelligent and flexible deep learning systems. \par

@@

\section{Embedded Meta-Learning (EML) architecture}
\vspace{-0.5em} % section
To address these challenges, we propose EML architectures, which are composed of two components: 
\vspace{-1em}
\begin{enumerate} \setlength \itemsep{0em}
\item An input/output system: domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$.
\item A meta-learning system which embeds tasks into the shared embedding space $Z$, and learns to map task embeddings to behavior.
\end{enumerate}
\vspace{-0.75em}
Why have a completely shared space $Z$ for data and tasks? The primary advantage is that it allows for arbitrary mappings between these entities. In addition to basic tasks, the system could in principle learn to perform meta-mappings to follow instructions or change behavior. \par
Without training on meta-mappings, of course, the system will not be able to execute them well. However, if it is trained on a broad enough set of such mappings, it will be able to generalize to new instances drawn from the same meta-task distribution, as we demonstrate below. For instances that fall outside its data distribution, or for optimal performance, it may require some retraining, however. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relatively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow \citep[cf.][]{Kumaran2016, Botvinick2019}. \par 
\begin{figure}
\centering
\begin{subfigure}[b]{0.72\textwidth}
\begin{tikzpicture}[auto, scale=0.66, every node/.style={scale=0.85}]
\node [dashblock] at (0.2, 0) (rep) {\begin{tabular}{c}$\bm{Z}$: Shared \\ representation\end{tabular}};

@@

\caption{The EML architecture and an inference step. (a) A schematic of the architecture. Blocks with solid edges denote learnable deep networks, dashed edges represent inputs, outputs, embeddings, etc., and $F$ is a deep network with parameters specified by $\mathcal{H}$. Note that there may be multiple input, output, and target decoders if the system has multiple input modalities, e.g. both language and vision. See appendix \ref{app_model_details} for details. (b) Depiction of a meta-learner forward pass. This process is identical between basic tasks and meta-mappings, except for where the example and probe embeddings come from, and how the loss or evaluation is computed. For basic inference, the example and probe embeddings come from basic inputs and outputs, and loss is computed in the output space. FOr meta-mapping, the example and probe input and outputs are task embeddings of basic tasks, and loss for training is computed in the embedding space, while evaluation is done by using the predicted embedding to perform the target task.} \label{architecture_inference_fig}

@@

Specifically, we embed inputs, targets, and mappings into a shared representational space $Z$ (see appendix \ref{app_lesion_results_shared_z} for discussion of why this space should be shared). Inputs are embedded by a deep network $\mathcal{I}: \text{input} \rightarrow Z$. Model outputs are decoded from $Z$ by a deep network $\mathcal{O}: Z \rightarrow \text{output}$. Target outputs are encoded by a deep network $\mathcal{T}: \text{targets} \rightarrow Z$.\par

@@

\paragraph{Meta-mapping:} However, the system is not limited to simply mapping inputs to outputs. The fundamental insight of our paper is to show how basic tasks and meta-mappings can be treated homogenously, by allowing the network to recurse on its own representations. From the perspective of our architecture, learning a meta-mapping between tasks is exactly analogous to learning a basic task. Anything that is embedded in $Z$ can be transformed using the same system. Because tasks are embedded in $Z$ for basic meta-learning, this allows for meta-mappings using exactly the same $\mathcal{M}$ and $\mathcal{H}$ that we use for basic tasks. For example, suppose we have an embedding $z_{game1} \in Z$ for the task of playing some game. We can generate a meta-mapping embedding $z_{\text{meta}} \in Z$ from embeddings generated by the system when it is trying to win and lose various games: $z_{\text{meta}} = \mathcal{M}\left( \left\{\left((z_{game2},z_{game2,lose}\right), ... \right\}\right)$. We can generate a new task embedding $\hat{z}_{new} \in Z$:  

>>>>

\begin{document}
\title{Zero-shot task adaptation by meta-mapping from task to task}
\author{%
Andrew K. Lampinen\\
Department of Psychology\\
Stanford University\\
\texttt{lampinen@stanford.edu}\\
\And
James L. McClelland\\
Department of Psychology\\
Stanford University\\
\texttt{mcclelland@stanford.edu}\\
}
\date{}
\maketitle

\begin{abstract}
How can deep learning systems flexibly reuse their knowledge? Toward this goal, we propose a new class of challenges, and a class of architectures that can solve them. The challenges are meta-mappings, which involve systematically transforming task behaviors to adapt to new tasks zero-shot. We suggest that the key to achieving these challenges is representing the task being performed along with the computations used to perform it. We therefore draw inspiration from functional programming and recent work in meta-learning to propose a class of Embedded Meta-Learning (EML) architectures that represent both data and tasks in a shared latent space. EML architectures are applicable to any type of machine learning task, including supervised learning and reinforcement learning. We demonstrate the utility of this perspective by exhibiting zero-shot remapping of behavior to adapt to new tasks.
\end{abstract}

\vspace{-0.75em} % section
\section{Introduction}
\vspace{-0.5em} % section
Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep[e.g.][]{Lake2016, Marcus2018}. One fundamental reason for this is that humans are aware of what we are trying to compute and why. By contrast, there is a fundamental separation of knowledge within most deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between data and tasks. That is, a neural network's knowledge about what is being computed is only implicitly accessible to those computations. \par
\looseness=-1
There are a number of advantages to representing knowledge about data and tasks together. In particular, it can grant the ability to rapidly adapt behavior to a new task. The problem of rapid learning has been partially addressed by meta-learning systems \citep[e.g.][see also section \ref{sec_discussion}]{Santoro2016, Finn2017a, Finn2018, Stadie2018, Botvinick2019}. However, humans can use our knowledge of a task to flexibly alter our behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance without any retraining (i.e. zero-shot). Deep learning systems at present generally lack this representational flexibility. \par
In this paper, we propose a new class of tasks based on this idea: meta-mappings, i.e. mappings between tasks (see below). This type of transfer is easily accessible to humans \citep{Lake2016}, but is generally inaccessible to most deep-learning models. To address this challenge, we propose using architectures which essentially take a functional perspective on meta-learning, and exploit the idea of homoiconicity. (A homoiconic programming language is one in which programs in the language can be manipulated by programs in the language, just as data can.) By treating both data and task behaviors as functions, we can conceptually think of both data \emph{and} learned task behaviors as transformable. This yields the ability to not only learn to solve new tasks, but to learn how to transform these solutions in response to changing task demands. We demonstrate that our architectures can flexibly remap their behavior to address the meta-mapping challenge. By allowing the network to recurse on its representations, our approach is able to achieve this flexibility parsimoniously, without adding additional parameters beyond what we use for basic meta-learning. We suggest that approaches like EML will be key to building more intelligent and flexible deep learning systems. \par

@@

\section{Embedded Meta-Learning (EML) architecture}
\vspace{-0.5em} % section
To address these challenges, we propose EML architectures, which are composed of two components: 
\vspace{-1em}
\begin{enumerate} \setlength \itemsep{0em}
\item An input/output system: domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$.
\item A meta-learning system which embeds tasks into the shared embedding space $Z$, and learns to map task embeddings to behavior.
\end{enumerate}
\vspace{-0.75em}
Why have a completely shared space $Z$ for individual data points and tasks? The primary advantage is that it allows for arbitrary mappings between these entities. In addition to basic tasks, the system could in principle learn to perform meta-mappings to follow instructions or change behavior. \par
Without training on meta-mappings, of course, the system will not be able to execute them well. However, if it is trained on a broad enough set of such mappings, it will be able to generalize to new instances drawn from the same meta-task distribution, as we demonstrate below. For instances that fall outside its data distribution, or for optimal performance, it may require some retraining, however. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relatively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow \citep[cf.][]{Kumaran2016, Botvinick2019}. \par 
\begin{figure}
\centering
\begin{subfigure}[b]{0.72\textwidth}
\begin{tikzpicture}[auto, scale=0.66, every node/.style={scale=0.85}]
\node [dashblock] at (0.2, 0) (rep) {\begin{tabular}{c}$\bm{Z}$: Shared \\ representation\end{tabular}};

@@

\caption{The EML architecture and an inference step. (a) A schematic of the architecture. Blocks with solid edges denote learnable deep networks, dashed edges represent inputs, outputs, embeddings, etc., and $F$ is a deep network with parameters specified by $\mathcal{H}$. Note that there may be multiple input, output, and target decoders if the system has multiple input modalities, e.g. both language and vision. See appendix \ref{app_model_details} for details. (b) Depiction of a meta-learner forward pass. This process is identical between basic tasks and meta-mappings, except for where the example and probe embeddings come from, and how the loss or evaluation is computed. For basic inference, the example and probe embeddings come from basic inputs and outputs, and loss is computed in the output space. For meta-mapping, the example and probe input and outputs are task embeddings of basic tasks, and loss for training is computed in the embedding space, while evaluation is done by using the predicted embedding to perform the target task.} \label{architecture_inference_fig}

@@

Specifically, we embed each input, target, or mapping into a shared representational space $Z$. This means that single data points are embedded in the same space as the representation of a functionr or an entire dataset. See appendix \ref{app_lesion_results_shared_z} for discussion of why $Z$ should be shared. Inputs are embedded by a deep network $\mathcal{I}: \text{input} \rightarrow Z$. Model outputs are decoded from $Z$ by a deep network $\mathcal{O}: Z \rightarrow \text{output}$. Target outputs are encoded by a deep network $\mathcal{T}: \text{targets} \rightarrow Z$.\par

@@

\paragraph{Meta-mapping:} However, the system is not limited to simply mapping inputs to outputs. The fundamental insight of our paper is to show how basic tasks and meta-mappings can be treated homogenously, by allowing the network to recurse on its own representations. From the perspective of our architecture, learning a meta-mapping between tasks is exactly analogous to learning a basic task. Anything that is embedded in $Z$ can be transformed using the same system. Because tasks are embedded in $Z$ for basic meta-learning, this allows for meta-mappings using exactly the same $\mathcal{M}$ and $\mathcal{H}$ that we use for basic tasks. Just as we would take a set of paired embeddings of data points for a basic task, and use them to compute a function embedding for that task, we can take a set of paired function embeddings, and use them to create an embedding for the meta-mapping. We can then use this meta-mapping embedding to transform another task. We can thus behave zero-shot on a novel task based on its relationship to a prior task. \par
For example, suppose we have an embedding $z_{game1} \in Z$ for the task of playing some game. We can generate a meta-mapping embedding $z_{\text{meta}} \in Z$ from embeddings generated by the system when it is trying to win and lose various games: $z_{\text{meta}} = \mathcal{M}\left( \left\{\left((z_{game2},z_{game2,lose}\right), ... \right\}\right)$. We can generate a new task embedding $\hat{z}_{new} \in Z$:  
